---
title: "Green_Buildings"
author: "Haritha Maheshkumar, Sachin Balakrishnan, Sahana Subramanian, Sijo VM"
date: "19/08/2018"
output: md_document
---

## **Visual story telling: Green buildings**

```{r, include=FALSE}
data_directory <- 'C:/Users/Sahana/Documents/GitHub/Data Visualization/Green Buildings'
setwd(data_directory)
data_raw <- read.csv('greenbuildings.csv')
```

```{r, include=FALSE}
library(corrplot)
library(LICORS)
library(RColorBrewer)
library(mosaic)
library(tidyverse)
```


```{r, include=FALSE}
data_raw$green_rating <- as.factor(data_raw$green_rating)
```

```{r, include=FALSE}
paste("Median rent for green buildings: ", 
median(data_raw$Rent[data_raw$green_rating == 1]))

paste("Median rent for non-green buildings: ", 
median(data_raw$Rent[data_raw$green_rating == 0]))

```

### **Problem**
An Austin real-estate developer is interested in the possible economic impact of "going green" in her latest project: a new 15-story mixed-use building on East Cesar Chavez, just across I-35 from downtown. Will investing in a green building be worth it, from an economic perspective? The baseline construction costs are $100 million, with a 5% expected premium for green certification.

The developer has had someone on her staff, who's been described to her as a "total Excel guru from his undergrad statistics course," run some numbers on this data set and make a preliminary recommendation. Here's how this person described his process:

> I began by cleaning the data a little bit. In particular, I noticed that a handful of the buildings in the data set had very low occupancy rates (less than 10% of available space occupied). I decided to remove these buildings from consideration, on the theory that these buildings might have something weird going on with them, and could potentially distort the analysis. Once I scrubbed these low-occupancy buildings from the data set, I looked at the green buildings and non-green buildings separately. The median market rent in the non-green buildings was $25 per square foot per year, while the median market rent in the green buildings was $27.60 per square foot per year: about $2.60 more per square foot. (I used the median rather than the mean, because there were still some outliers in the data, and the median is a lot more robust to outliers.) Because our building would be 250,000 square feet, this would translate into an additional $250000 x 2.6 = $650000 of extra revenue per year if we build the green building.

> Our expected baseline construction costs are $100 million, with a 5% expected premium for green certification. Thus we should expect to spend an extra $5 million on the green building. Based on the extra revenue we would make, we would recuperate these costs in $5000000/650000 = 7.7 years. Even if our occupancy rate were only 90%, we would still recuperate the costs in a little over 8 years. Thus from year 9 onwards, we would be making an extra $650,000 per year in profit. Since the building will be earning rents for 30 years or more, it seems like a good financial move to build the green building.

> Goal: The developer listened to this recommendation, understood the analysis, and still felt unconvinced. She has therefore asked you to revisit the report, so that she can get a second opinion.

### **Visualizations**

```{r ,out.width=c('50%', '50%'), fig.show='hold', echo=FALSE}
ggplot(data=data_raw) + 
  geom_point(mapping=aes(x=cluster_rent, y=Rent, colour=green_rating)) +
  labs(x="Cluster Rent", y='Rent', title = 'Green buildings: Cluster Rent VS Rent',
       color='Green building')

ggplot(data=data_raw) + 
  geom_point(mapping=aes(x=age, y=Rent, colour=green_rating))+
  labs(x="Age", y='Rent', title = 'Green buildings: Age VS Rent',
       color='Green building')

ggplot(data=data_raw) + 
  geom_point(mapping=aes(x=size, y=Rent, colour=green_rating)) +
  labs(x="Size", y='Rent', title = 'Green buildings: Size VS Rent',
       color='Green building')

ggplot(data=data_raw) + 
  geom_point(mapping=aes(x=leasing_rate, y=Rent, colour=green_rating)) +
  labs(x="Leasing Rate", y='Rent', title = 'Green buildings: Leasing Rate VS Rent',
       color='Green building')


ggplot(data=data_raw) + 
  geom_point(mapping=aes(x=age, y=Rent, colour=class_a))+
  labs(x="Age", y='Rent', title = 'Class A: Age VS Rent',
       color='Class A building')

```

**Observations** 

* Rent is correlated with the cluster rent
* Rent is correlated with the size, as expected
* Most of the class A buildings are also younger 
* Age does not have a high correlation with rent 
* Class a buildings get higher rent as they are premium buildings

```{r out.width=c('50%', '50%'), fig.show='hold', echo=FALSE}
g = ggplot(data_raw, aes(x=age))
g + geom_density(aes(fill=factor(green_rating)), alpha=0.4)+
  labs(x="Age", y='Density', title = 'Distribution of age',
       fill='Green building')

ggplot(data_raw, aes(class_a, ..count..)) + geom_bar(aes(fill = green_rating), position = "dodge")+
  labs(x="Class a", y='Number of buildings', title = 'Class A vs Green Buildings',
       fill='Green building')

g = ggplot(data_raw, aes(x=size))
g + geom_density(aes(fill=factor(green_rating)), alpha=0.4)+
  labs(x="Size", y='Density', title = 'Distribution of size',
       fill='Green building')


medians <- aggregate(Rent ~  class_a, data_raw, median)
ggplot(data=data_raw, aes(x=factor(class_a), y=Rent, fill=class_a)) + geom_boxplot()+
  stat_summary(fun.y=median, colour="darkred", geom="point", 
               shape=18, size=3,show.legend = FALSE) + 
  geom_text(data = medians, aes(label = Rent, y = Rent - 20)) +
  labs(x="Class A", y='Rent', title = 'Rent vs Class a',
       fill='Class A')

```

**Observations** 

* Most of the green buildings are younger than non-green buildings 
* The proportion of class a buildings is higher in green buildings 
* The proportion of green and non-green building increases as the size of buildings increases
* The is a significant difference in the of rent of class a and non-class a buildings 

```{r out.width=c('50%', '50%'), fig.show='hold', echo=FALSE}
data_raw$age_cat <- cut(data_raw$age, breaks = c(0, seq(10, 190, by = 10)), labels = 0:18,right=FALSE)

medians <- aggregate(Rent~ age_cat + green_rating, data_raw, median)


ggplot(data = medians, mapping = aes(y = Rent, x = age_cat ,group = green_rating, colour=green_rating)) +
   geom_line(size=1.2) +
  labs(x="Age in 10 years", y='Median Rent', title = 'All buildings: Median rent over the years',
       fill='Green building')

# Size in 100k
data_raw$size_cat <- cut(data_raw$size, breaks = c(0, seq(10, 3781045, by = 100000)), labels = 0:37,right=FALSE)
medians <- aggregate(Rent ~ size_cat + green_rating, data_raw, median)
ggplot(data = medians, mapping = aes(y = Rent, x = size_cat ,group = green_rating, colour=green_rating)) +
   geom_line(size=1.2) +
  labs(x="Size in 100k sq.ft", y='Median Rent', title = 'All buildings: median rent for different building sizes',
       fill='Green building')


data_non_class_a <- subset(data_raw, data_raw$class_a != 1)
data_non_class_a$age_cat <- cut(data_non_class_a$age, breaks = c(0, seq(10, 190, by = 10)), labels = 0:18,right=FALSE)

medians <- aggregate(Rent~ age_cat + green_rating, data_non_class_a, median)


ggplot(data = medians, mapping = aes(y = Rent, x = age_cat ,group = green_rating, colour=green_rating)) +
   geom_line(size=1.2)+
  labs(x="Age in 10 years", y='Median Rent', title = 'Non-Class A buildings: Median rent over the years',
       fill='Green building')

# Size in 100k
data_non_class_a$size_cat <- cut(data_non_class_a$size, breaks = c(0, seq(10, 3781045, by = 100000)), labels = 0:37,right=FALSE)
medians <- aggregate(Rent ~ size_cat + green_rating, data_non_class_a, median)
ggplot(data = medians, mapping = aes(y = Rent, x = size_cat ,group = green_rating, colour=green_rating)) +
   geom_line(size=1.2)+
  labs(x="Size in 100k sq.ft", y='Median Rent', title = 'Non-class A buildings: median rent for different building sizes',
       fill='Green building')


```


**Observations** 

* For a size of 250,000 sqft, the green buildings have a higher rent when it is a class a building
* The rent of green buildings is lower than non-green ones when they are not class a buildings
* The rent difference is not uniform across different sizes and ages 




```{r, echo=FALSE, include=FALSE}
data_size <- subset(data_raw, data_raw$size > 200000 & data_raw$size < 300000)
data_size <- subset(data_size, data_size$class_a == 1)
data_size_class <- subset(data_raw, data_non_class_a$size > 200000 & data_non_class_a$size < 300000)
paste("Median leasing rate for class a buildings of sizes ranging from 200k to 300k sq.ft ", 
median(data_size$leasing_rate))

medians <- aggregate(Rent~ age_cat + green_rating, data_size, median)
medians_1 <- subset(medians, medians$green_rating == 1)
rent_1<-medians_1[1:5,]$Rent
medians_0 <- subset(medians, medians$green_rating == 0)
rent_0<-medians_0[1:5,]$Rent

paste("Difference in rent for the first 5 years class a buildings: ", 
(sum(rent_1,na.rm = T) - sum(rent_0, na.rm = T)) / 5)


medians <- aggregate(Rent~ age_cat + green_rating, data_size_class, median)
medians_1 <- subset(medians, medians$green_rating == 1)
rent_1<-medians_1[1:5,]$Rent
medians_0 <- subset(medians, medians$green_rating == 0)
rent_0<-medians_0[1:5,]$Rent

paste("Difference in rent for the first 5 years for non-class a buildings: ", 
(sum(rent_1,na.rm = T) - sum(rent_0, na.rm = T)) / 5 )
```

### **Insights and Recommendations** 

We have seen that the analysis by stats guru is flawed since he fails to account for all the factors that affect the rent. First, he used the median rent of all buildings to calculate the returns. Hence he fails to accout for other factors such as size and class of the buildings into his analysis. For instance, we have
seen that class a buildings will yield a higher rent than non green buildings that are class a. 

**Calculations**  

* The rent difference is not uniform across different sizes and age, so we cannot use a fixed difference in rent to calculate the returns
* For the same reason, we should consider only the buildings that have sizes between 200k and 300k sq.ft 
* We should also use the median leasing rate of such buildings instead of 90% rate 
* The data provided does not have information about class a buildings with sizes ranging from 200k sq.ft to 300k sq.ft. So let's use average 5 year return to arrive at final recommendations



```{r, include=FALSE}

paste("If we build a class a green building and if we assume 91.6% occpancy rate, it is expected to recuperate the costs in  ", round(5000000/(3.097*250000*0.916),2), " years")

```
**Final recommendation** 

* If the building is not a Class-A building, it is not wise to invest in a green building since the average returns for 5 years are negative
* The builder should invest in a Class-A green building to yield positive returns
* We can expect a occupancy rate of 91.6% on such buildings 
* The average difference in rent for green and non-green buildings that are class a and whose sizes ranging from 200k to 300k is 3.097
* Hence, for a 250k sq.ft building at 91.6% occupancy, we expect to recuperate the costs in 7.05 years





