---
title: "Portfolio_Modeling"
author: "Haritha Maheshkumar, Sachin Balakrishnan, Sahana Subramanian, Sijo VM"
date: "17/01/2020"
output: md_document
---
## **Portfolio Modeling**

### **Problem**
In this problem, we will construct three different portfolios of exchange-traded funds, or ETFs, and use bootstrap resampling to analyze the short-term tail risk of our portfolios.


```{r, echo=FALSE, include=FALSE}
library(ggstance)
library(mosaic)
library(quantmod)
library(foreach)

```




We selected the ETFs ensuring diversity and different levels of risk. 

The Ivesco QQQ trust is one of the largest, owns only non-financial stocks and is tech-heavy. The stock had performed well in 2017 but had a poor return in 2018. 

SPY is one of the safest and largest ETFs around.

ProShares VIX Short-Term Futures ETF (SVXY) is a high risk ETF. This is an unusual ETF since the performance is dependent on the market volatility, not security. 

ProShares UltraShort FTSE Europe (EPV) is a low performing ETF for the past few years. iShares Core Growth Allocation ETF (AOR) is a very diverse ETF. 

In total, we have selected 6 ETFs - "QQQ", "SPY", "SVXY", "EPV", "AOR" and "YYY". We have considered 5 years of ETF data starting from 01-Jan-2014.



```{r, echo=FALSE, include=FALSE}

# Import a few stocks
mystocks = c("QQQ", "SPY", "SVXY", "EPV", "AOR", "YYY")

# Getting the price data for 5 years
getSymbols(mystocks, from='2014-01-01')

for(ticker in mystocks){
  expr = paste0(ticker, "a=adjustOHLC(", ticker, ")")
  eval(parse(text=expr))
}


```


Sample Data for YYY 

```{r, echo=FALSE}
head(YYYa)

```



```{r, echo=FALSE, include=FALSE}

# Computing the return matrix

all_returns = cbind( ClCl(QQQa),
                     ClCl(SPYa),
                     ClCl(SVXYa),
                     ClCl(EPVa),
                     ClCl(AORa),
                     ClCl(YYYa))

head(all_returns)

```



```{r, echo=FALSE, include=FALSE}
# Remove NAs

all_returns = as.matrix(na.omit(all_returns))
head(all_returns)

```



Lets look at how the stocks are performing relative to each other. We can see a strong correlation here. But it is complex and non-linear. As discussed above, a few are performing well, others  are not.

```{r, echo=FALSE}

# checking the correlation
pairs(all_returns)

```


Volatility of the ETFs across the 5 year period.


```{r, echo=FALSE}

# Volatility check
plot(ClCl(QQQa), type='l')
plot(ClCl(SPYa), type='l')
plot(ClCl(SVXYa), type='l')
plot(ClCl(EPVa), type='l')
plot(ClCl(AORa), type='l')
plot(ClCl(YYYa), type='l')

```





```{r, echo=FALSE, include=FALSE}

initial_wealth = 100000


```



Our initial wealth is $100,000 

SIMULATION 1 : SAFE Portfolio

ETFs: "QQQ", "SPY", "SVXY", "EPV", "AOR", "YYY"

For the safe portfolio, we distributed 90% of the total wealth among the high performing ETFs - QQQ, SPY and AOR.


```{r, echo=FALSE, include=FALSE}

sim1 = foreach(i=1:5000, .combine = rbind) %do% {
  weights = c(0.3, 0.4, 0.03, 0.03, 0.2, 0.04)
  total_wealth = initial_wealth
  holdings = total_wealth * weights
  n_days = 20
  wealthtracker = rep(0, n_days)
  
  for(today in 1:n_days){
    return_today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings * (1 + return_today)
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
    
    # Rebalancing
    holdings = total_wealth * weights
  }
  
  wealthtracker
}

head(sim1)
hist(sim1[,n_days], 50)
plot(density(sim1[,n_days]))

# Profit/loss

hist(sim1[,n_days]- initial_wealth, breaks=30)
conf_5Per = confint(sim1[,n_days]- initial_wealth, level = 0.90)$'5%'
cat('\nAverage return of investement after 20 days', mean(sim1[,n_days]), "\n")
cat('\n5% Value at Risk for safe portfolio-',conf_5Per, "\n")

```

```{r, echo=FALSE, include=FALSE}
wealth_daywise = c()
  
for (i in 1:n_days){
    wealth_daywise[i] = mean(sim1[,i]) 
}

days = 1:n_days
df = data.frame(wealth_daywise, days)

```


```{r, echo=FALSE}
ggplot(data=df, aes(x=days, y=wealth_daywise, group=1)) +
  geom_line(color="red")+
  geom_point() +
  xlab('Days') +
  ylab('Return of investments') + 
  ggtitle('Safe Portfolio: Retruns over 20 days')
```



```{r, echo=FALSE}

hist(sim1[,n_days], 50)
plot(density(sim1[,n_days]))
hist(sim1[,n_days]- initial_wealth, breaks=30)
conf_5Per = confint(sim1[,n_days]- initial_wealth, level = 0.90)$'5%'
cat('\nAverage return of investement after 20 days', mean(sim1[,n_days]), "\n")
cat('\n5% Value at Risk for safe portfolio-',conf_5Per, "\n")

```



SIMULATION 2 : HIGH RISK PORTFOLIO

For the high risk portfolio, we distributed 90% of the total wealth among the low performing ETFs - SVXY, EPV and YYY.

Average return of investement after 20 days - $100724.3  
5% Value at Risk for safe portfolio - $7850.127  

```{r, echo=FALSE, include=FALSE}

sim2 = foreach(i=1:5000, .combine = rbind) %do% {
  weights = c(0.01, 0.02, 0.3, 0.3, 0.07, 0.3)
  total_wealth = initial_wealth
  holdings = total_wealth * weights
  n_days = 20
  wealthtracker = rep(0, n_days)
  
  for(today in 1:n_days){
    
    return_today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings * (1 + return_today)
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
    
    # Rebalancing
    holdings = total_wealth * weights
  }
  
  wealthtracker
}

head(sim2)
hist(sim2[,n_days], 50)
plot(density(sim2[,n_days]))

# Profit/loss
hist(sim2[,n_days]- initial_wealth, breaks=30)

hist(sim2[,n_days]- initial_wealth, breaks=30)
conf_5Per = confint(sim2[,n_days]- initial_wealth, level = 0.90)$'5%'
cat('\nAverage return of investement after 20 days', mean(sim2[,n_days]), "\n")
cat('\n5% Value at Risk for High portfolio-',conf_5Per, "\n")

```


```{r, echo=FALSE, include=FALSE}
wealth_daywise = c()
  
for (i in 1:n_days){
    wealth_daywise[i] = mean(sim2[,i]) 
}

days = 1:n_days
df = data.frame(wealth_daywise, days)

```


```{r, echo=FALSE}

ggplot(data=df, aes(x=days, y=wealth_daywise, group=1)) +
  geom_line(color="red")+
  geom_point() +
  xlab('Days') +
  ylab('Return of investments') + 
  ggtitle('High Risk Portfolio: Retruns over 20 days')

```

```{r, echo=FALSE}

hist(sim2[,n_days], 50)
plot(density(sim2[,n_days]))

# Profit/loss
hist(sim2[,n_days]- initial_wealth, breaks=30)
conf_5Per = confint(sim2[,n_days]- initial_wealth, level = 0.90)$'5%'
cat('\nAverage return of investement after 20 days', mean(sim2[,n_days]), "\n")
cat('\n5% Value at Risk for High portfolio-',conf_5Per, "\n")
```



SIMULATION 3 - With equal weights for high risk and low risk

```{r, echo=FALSE, include=FALSE}
sim3 = foreach(i=1:5000, .combine = rbind) %do% {
  weights = c(0.17, 0.17, 0.17, 0.17, 0.17, 0.15)
  total_wealth = initial_wealth
  holdings = total_wealth * weights
  n_days = 20
  wealthtracker = rep(0, n_days)
  
  for(today in 1:n_days){
    
    return_today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings * (1 + return_today)
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
    
    # Rebalancing
    holdings = total_wealth * weights
  }
  
  wealthtracker
}

head(sim3)
```



```{r, echo=FALSE}
hist(sim3[,n_days], 50)
plot(density(sim3[,n_days]))

# Profit/loss
hist(sim3[,n_days]- initial_wealth, breaks=30)
conf_5Per = confint(sim3[,n_days]- initial_wealth, level = 0.90)$'5%'
cat('\nAverage return of investement after 20 days', mean(sim3[,n_days]), "\n")
cat('\n5% Value at Risk for High portfolio-',conf_5Per, "\n")
```

```{r, echo=FALSE, include=FALSE}

wealth_daywise = c()
for (i in 1:n_days){
    wealth_daywise[i] = mean(sim3[,i]) 
}

days = 1:n_days
df = data.frame(wealth_daywise, days)

```


```{r, echo=FALSE}

ggplot(data=df, aes(x=days, y=wealth_daywise, group=1)) +
  geom_line(color="red")+
  geom_point() +
  xlab('Days') +
  ylab('Return of investments') + 
  ggtitle('Diverse Portfolio: Retruns over 20 days')

```

#### **Summary**

For the safe portfolio, we are observing the maximum return of investment and the lowest 5% VaR. 
As the portfolio risk increases, we are able to witness the decrease in returns and increase in VaR value as expected.

References:
https://www.bankrate.com/investing/best-etfs/
https://etfdb.com/compare/lowest-ytd-returns/
