---
title: "Association Rule Mining"
author: "Haritha Maheshkumar, Sachin Balakrishnan, Sahana Subramanian, Sijo VM"
date: "17/01/2020"
output: md_document
---
## **Association rule mining**

```{r echo=FALSE, include=FALSE}
## Load the required packages
library(tidyverse)
library(arules) 
library(arulesViz)
```

### **Problem**
We have multiple shopping baskets of grocery purchases and we will find some interesting association rules between then using rule mining.

#### Presenting the structure of the raw dataset:

```{r echo=FALSE}
## Read in the dataset and explore the structure
setwd("C:/Users/Sahana/Documents/GitHub/Data-Science-Mini-Projects/Association Rule Mining")
groceries_raw = scan("groceries.txt", what = "", sep = "\n")
head(groceries_raw)
```

```{r echo=FALSE, include=FALSE}
str(groceries_raw)
summary(groceries_raw)
```

We transform the data into a "transactions" class before applying the apriori algorithm.
The summary of the dataset reveals the following:
1. There are total of 9835 transactions in our dataset
2. Whole milk is the present in 2513 baskets and is the most frequently bought item
3. More than half of the transactions have 4 or lesser items per basket

```{r echo=FALSE, include=FALSE}
## Process the data and cast it as a "transactions" class
groceries = strsplit(groceries_raw, ",")
groctrans = as(groceries, "transactions")
summary(groctrans)
```

```{r echo=FALSE}
itemFrequencyPlot(groctrans, topN = 20)
```

#### **Let's explore rules with support > 0.05, confidence > 0.1 and length <= 2 using the 'apriori' algorithm**
There are only 6 rules generated because of the high support and low confidence level. We also notice that most relationships in this item set include whole milk, yogurt and rolls/buns which is in accordance with the transaction frequency plot we saw earlier. These are some of the most frequently bought items.

```{r echo=FALSE, include=FALSE}
grocrules_1 = apriori(groctrans, 
                     parameter=list(support=0.05, confidence=.1, minlen=2))
```

```{r echo=FALSE}
arules::inspect(grocrules_1)
plot(grocrules_1, method='graph')
```

#### **Let's decrease support further and increase confidence slightly with support > 0.02, confidence > 0.2 and length <= 2**

This item set contains 72 rules and includes a lot more items. However, whole milk still seems to be a common occurence.

```{r echo=FALSE, include=FALSE}
grocrules_2 = apriori(groctrans, 
                     parameter=list(support=0.02, confidence=.2, minlen=2))
arules::inspect(grocrules_2)
```

```{r echo=FALSE}
plot(head(grocrules_2,15,by='lift'), method='graph')
```


#### **Let us increase the confidence level and decrease the support further. Let's explore rules with support > 0.0015, confidence > 0.8 and length <= 2**


```{r echo=FALSE, include=FALSE}
grocrules_3 = apriori(groctrans, 
                     parameter=list(support=0.0015, confidence=0.8, minlen=2))
arules::inspect(grocrules_3)
```

```{r echo=FALSE}
plot(head(grocrules_3, 5, by='lift'), method='graph')
```

### **Summary**
From the association rules, some of the conclusions that can be drawn are:
1. People are more likely to buy bottled beer if they purchased red wine or liquor
2. People are more likely to buy vegetables when they buy vegetable/fruit juice
3. Whole milk is the most common item purchased by customers


